Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-19 article.tex
Update article.tex

Under that assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{koller2009probabilistic}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j | X^{-i,j}} \neq 0$,
where $\rho_{X_i, X_j | X^{-i,j}}$ can now be interpreted as the degree of association
between the two variables (i.e., the larger the value, the stronger the interaction).
Together, the partial correlation statistics correspond to the inverse
$\Sigma^{-1}$ of the covariance matrix, also known as the concentration of
precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our experiments, $M$ is set
to $0.8 p$, thereby allowing us to further filter noise in the data.

>>>>

Under that assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{koller2009probabilistic}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j | X^{-i,j}} \neq 0$,
where $\rho_{X_i, X_j | X^{-i,j}}$ can now be interpreted as the degree of association
between the two variables (i.e., the larger the value, the stronger the interaction).
Together, the partial correlation statistics correspond to the inverse
$\Sigma^{-1}$ of the covariance matrix, also known as the concentration or
precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our experiments, $M$ is set
to $0.8 p$, thereby allowing us to further filter noise in the data.
